---
title: "Programming with CLIFF and TIFF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Programming with CLIFF and TIFF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Note: this vignette requires `TIFF` and `CLIFF` to be installed.*

```{r, include = FALSE}

isTiffInstalled <- nzchar(system.file(package = "TIFF"))
isCliffInstalled <- nzchar(system.file(package = "CLIFF"))

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = isTiffInstalled && isCliffInstalled
)
```

```{r setup}
library(XIFF)

getXIFFOptions <- function() {
  glue::glue("Column: {getOption('xiff.column')},",
             " Schema: {getOption('xiff.schema')}")
}

XIFF::setDbOptions()
getXIFFOptions() # read options

# Call function from TIFF
geneAnno <- TIFF::getGeneAnno("human")
## read options
## they are the options from TIFF
## TIFF overwritten XIFF defaults
getXIFFOptions() 

# Call function from CLIFF
genAnnoCLIFF <- CLIFF::getGeneAnno()

## read options
## they are the options from CLIFF
## meaning that CLIFF overwritten
## options set by cliff
getXIFFOptions()

```
The behaviour as presented above might pose serious risks to the validity of the scripts that contains both `TIFF` and `CLIFF` code. What's even more, when loading both `CLIFF` and `TIFF` there are multiple function name clashes:

```r
Attaching package: ‘TIFF’

The following objects are masked from ‘package:CLIFF’:

    app, appUI_main, appUI_sidebar, differentialGeneExpression_LimmaVoom, differentialGeneSignature,
    differentialHallmarkSets, differentialProteinExpression, generateCopynumberPlot,
    generateCopynumberWaterfallPlot, generateExpressionHeatmap, generateExpressionPlot,
```

To solve the namespace and options issue two approaches are provided:

### Based on the `withr` package

The package specific code is evaluated in 
special function that takes care about setting the proper options and namespace.

```{r, message=FALSE, warning=FALSE}
library(CLIFF)
library(TIFF)

withCLIFF({
  print(getOption("xiff.column"))    # "celllinename"
  print(environment(getInformation)) # <environment: namespace:CLIFF>
  y <- getOption("xiff.column")
})

print(y) # value from withCLIFF

withTIFF({
  print(getOption("xiff.column"))        # "tissuename"
  print(environment(getInformation)) # <environment: namespace:TIFF>
  y <- getOption("xiff.column")
})

print(y) # value from withTIFF

```

### By setting the global options using `setCLIFF` and `setTIFF`

Note that for package specific functions using the `::` is recommended.

```{r}
setCLIFF()

### CLIFF code goes here ###
print(getOption("xiff.column"))
print(environment(CLIFF::getInformation))

setTIFF()

### TIFF code goes here ###
print(getOption("xiff.column"))
print(environment(TIFF::getInformation))

```

