---
title: "Feature Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Feature Selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Quick Intro

Currently, the `XIFF` package implements two functions for feature selection:
- `selectBestFeaturesBoruta` which uses `Boruta` package.
  - see e.g. https://cran.r-project.org/web/packages/Boruta/vignettes/inahurry.pdf for more details.
- `selectBestFeaturesTTest` which uses `genefilter::colttests`.
  - run `?genefilter::colttests` for more details.

The purpose of the example below is to show the basic API and execution time on some random data set.

```{r setup}
library(XIFF)
library(dplyr)
library(glue)

# Utility function that shows the run time for passed expression
test_time <- function(expr) {
  cr <- match.call()
  time <- system.time(res <- eval(expr))
  message("Running ", cr, " took ", round(time[[3]], 2), "seconds")
  res
}

data("feature_selection_data", package = "XIFF")

# Some very basic example
message(glue("Number of features: {ncol(feature_selection_data)}"))
featureFit <- test_time(selectBestFeaturesBoruta(feature_selection_data))
featureFitTTest <- test_time(selectBestFeaturesTTest(feature_selection_data))

message(glue("Number of features selected by Boruta: {ncol(featureFit$df)-1}"))
message(glue("Number of features selected by ColTest: {ncol(featureFitTTest$df)-1}"))
```

```{r}
# Plot the result
plot(featureFit$fit, las = 1, horizontal = TRUE)
# Show resulting table
featureFit$stats
```
###  Limit the number of features to be selected

```{r}
featureFit <- test_time(selectBestFeaturesBoruta(feature_selection_data, maxFeatures = 5))
plot(featureFit$fit, las = 1, horizontal = TRUE)
featureFitTTest <- test_time(selectBestFeaturesTTest(feature_selection_data, maxFeatures = 5))

s1 <- featureFit$stats %>% select(ensg) %>% mutate(BorutaSelection = ensg)
s2 <- featureFitTTest$stats %>% select(ensg) %>% mutate(TTestSelection = ensg)

# Comparsion of the selected variables:
full_join(s1, s2, by = "ensg")
```

# Notes:

- `selectBestFeaturesBoruta` algorithm is much slower than `selectBestFeaturesTTest`.
- The resulting variables are different.
